---
title: "Final Report"
output: html_document
date: "2024-05-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

- Explore airbnb data in chicago and determine factors that coorelate and produce higher rental rates, prices, and renter reviews. 

Here are my main questions I want to focus on

What are the best type of properties to own

Best room types to own

How many Accommodies in our property is Ideal

Initially I wanted to focus on marketing variables aswell, however after analysis its clear and straightforward that most serious airbnb
property listers use all marketing techniques the same way.


## Data
# structure

My data came from kaggle in a zip file containing an excel file. I opened Excel and fixed some issues involving variation in data entry format.
I wanted to focus on only chicago and only the most popular properties, for example: excluding boat rentals and log cabins etc.
I filtered the data to narrow the results of my dataset down.

I also removed outliers through IQR 

# cleaning

```{r}
airbnb_data <- read.csv("Airbnb_Data_ChrisDraper.csv")
head(airbnb_data)  
str(airbnb_data)   

airbnb_data$first_review <- as.Date(airbnb_data$first_review, format = "%m/%d/%Y")
airbnb_data$last_review <- as.Date(airbnb_data$last_review, format = "%m/%d/%Y")
airbnb_data$host_since <- as.Date(airbnb_data$host_since, format = "%m/%d/%Y")

airbnb_data$host_has_profile_pic <- ifelse(airbnb_data$host_has_profile_pic == "t", 1, 0)

table(airbnb_data$host_has_profile_pic)

airbnb_data$host_identity_verified <- ifelse(airbnb_data$host_identity_verified == "t", 1, 0)

table(airbnb_data$host_identity_verified)  

airbnb_data$cleaning_fee_numeric <- as.numeric(airbnb_data$cleaning_fee)

table(airbnb_data$cleaning_fee_numeric)  

airbnb_data$instant_bookable <- ifelse(airbnb_data$instant_bookable == "t", TRUE, FALSE)
airbnb_data$instant_bookable_numeric <- as.numeric(airbnb_data$instant_bookable)
table(airbnb_data$instant_bookable_numeric)

airbnb_data$X <- NULL

airbnb_data$host_response_rate <- as.numeric(gsub("%", "", airbnb_data$host_response_rate))

summary(airbnb_data)

airbnb_data$property_type <- factor(airbnb_data$property_type)

str(airbnb_data$property_type)

airbnb_data$room_type <- factor(airbnb_data$room_type)

str(airbnb_data$room_type)
levels(airbnb_data$room_type)

airbnb_data$bed_type <- factor(airbnb_data$bed_type)
str(airbnb_data$bed_type)
levels(airbnb_data$bed_type)


levels(airbnb_data$cancellation_policy)[levels(airbnb_data$cancellation_policy) == "super_strict_30"] <- "super_strict"
levels(airbnb_data$cancellation_policy)[levels(airbnb_data$cancellation_policy) == "super_strict_60"] <- "super_strict"
table(airbnb_data$cancellation_policy)

airbnb_data$cancellation_policy <- factor(airbnb_data$cancellation_policy)
str(airbnb_data$cancellation_policy)
levels(airbnb_data$cancellation_policy)

airbnb_data$city <- factor(airbnb_data$city)
str(airbnb_data$city)
levels(airbnb_data$city)

property_type_counts <- table(airbnb_data$property_type)
property_type_counts <- sort(property_type_counts, decreasing = TRUE) 

top_property_types <- names(property_type_counts[1:5])
print(top_property_types)

###########################################################

chicago_data <- airbnb_data[airbnb_data$city == "Chicago", ]

str(chicago_data)

summary(chicago_data)

#######################

remove_outliers <- function(df, column_name) {
  Q1 <- quantile(df[[column_name]], 0.25, na.rm = TRUE)
  Q3 <- quantile(df[[column_name]], 0.75, na.rm = TRUE)
  IQR <- Q3 - Q1
  lower_bound <- Q1 - 1.5 * IQR
  upper_bound <- Q3 + 1.5 * IQR
  filtered_data <- df[df[[column_name]] >= lower_bound & df[[column_name]] <= upper_bound, ]
  return(filtered_data)
}

chicago_data_filtered <- remove_outliers(chicago_data, "log_price")
chicago_data_filtered <- remove_outliers(chicago_data_filtered, "host_response_rate")
chicago_data_filtered <- remove_outliers(chicago_data_filtered, "number_of_reviews")
chicago_data_filtered <- remove_outliers(chicago_data_filtered, "review_scores_rating")

summary(chicago_data_filtered[c("log_price", "host_response_rate", "number_of_reviews", "review_scores_rating")])
cat("Number of observations after removing outliers:", nrow(chicago_data_filtered), "\n")

###################################################################

chicago_data_filtered$combined_score <- log1p(chicago_data_filtered$number_of_reviews) * chicago_data_filtered$review_scores_rating

####################################################################
chicago_data_filtered <- chicago_data_filtered[chicago_data_filtered$property_type %in% top_property_types, ]

chicago_data_filtered$property_type <- factor(chicago_data_filtered$property_type)

levels(chicago_data_filtered$property_type)
########################################################################

```

# variables

I changed many variables to factors and included boolean values as integers ranging from 0 - 1. 0 being false and 1 being true.
I continued filtering data throughtout my analysis to achieve grpahs that visually showed what I was trying to represent.
I filtered down to the top 5 most rented property_types 

I also created a new metric to weigh the Renting frequency and renters score variables. I multiplied number of reviews and average review score together to obtain a combined score metric that weighted both the variables into a more fair mertic that I will use to conduct further analysis



```{r}
########################################################################
library(ggplot2)
library(dplyr)

ggplot(chicago_data_filtered, aes(x = property_type, y = combined_score, fill = property_type)) +
  geom_boxplot() +
  labs(title = "Box Plot of Combined Scores by Property Type in Chicago",
       x = "Property Type",
       y = "Combined Score") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") 


#####################################################################


ggplot(chicago_data_filtered, aes(x = room_type, y = combined_score, fill = room_type)) +
  geom_boxplot() +
  labs(title = "Box Plot of Combined Scores by Room Type in Chicago",
       x = "Room Type",
       y = "Combined Score") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")

####################################################################

room_property_table <- table(chicago_data_filtered$room_type, chicago_data_filtered$property_type)

print(room_property_table)

ggplot(data = as.data.frame(room_property_table), aes(x = Var2, y = Freq, fill = Var1)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Count of Room Types by Property Type",
       x = "Property Type",
       y = "Count",
       fill = "Room Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

###################################################################

ggplot(chicago_data_filtered, aes(x = combined_score, y = log_price)) +
  geom_point(aes(color = room_type), alpha = 0.5) +
  labs(title = "Scatter Plot of Log Price by Combined Score",
       x = "Combined Score",
       y = "Log Price",
       color = "Room Type") +
  theme_minimal() +
  theme(legend.position = "right")

###################################################################

filtered_data <- chicago_data_filtered %>%
  filter(room_type == "Entire home/apt", 
         property_type %in% c("Apartment", "Condominium", "House"))

mean_scores_by_accommodates <- filtered_data %>%
  group_by(accommodates) %>%
  summarise(mean_combined_score = mean(combined_score, na.rm = TRUE))

ggplot(mean_scores_by_accommodates, aes(x = factor(accommodates), y = mean_combined_score, fill = factor(accommodates))) +
  geom_bar(stat = "identity") +
  labs(title = "Mean Combined Score by Accommodates for Entire Homes/Apts",
       subtitle = "Filtered by Property Types: Apartment, Condominium, House",
       x = "Accommodates",
       y = "Mean Combined Score") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")

####################################################################


filtered_data <- chicago_data_filtered %>%
  filter(room_type == "Entire home/apt", 
         property_type %in% c("Apartment", "Condominium", "House"))

ggplot(filtered_data, aes(x = combined_score, y = accommodates)) +
  geom_point(aes(color = property_type), alpha = 0.6) +
  labs(title = "Scatter Plot of Combined Score vs. Accommodates",
       subtitle = "Filtered by Property Types: Apartment, Condominium, House",
       x = "Combined Score",
       y = "Accommodates") +
  scale_color_manual(values = c("Apartment" = "blue", "Condominium" = "green", "House" = "red")) +
  theme_minimal() +
  theme(legend.title = element_blank(),
        legend.position = "bottom")

#######################################################################

filtered_data <- chicago_data_filtered %>%
  filter(room_type == "Entire home/apt", 
         property_type %in% c("Apartment", "Condominium", "House"))

ggplot(filtered_data, aes(x = log_price, y = accommodates)) +
  geom_point(aes(color = property_type), alpha = 0.6) +
  labs(title = "Scatter Plot of log_price vs. Accommodates",
       subtitle = "Filtered by Property Types: Apartment, Condominium, House",
       x = "log_price",
       y = "Accommodates") +
  scale_color_manual(values = c("Apartment" = "blue", "Condominium" = "green", "House" = "red")) +
  theme_minimal() +
  theme(legend.title = element_blank(),
        legend.position = "bottom")

########################################################################

filtered_data <- chicago_data_filtered %>%
  filter(room_type == "Entire home/apt", 
         property_type %in% c("Apartment", "Condominium", "House"))

ggplot(filtered_data, aes(x = log_price, y = accommodates)) +
  geom_point(aes(color = property_type), alpha = 0.6) +  
  geom_smooth(method = "lm", se = TRUE, color = "black") +  
  labs(title = "Scatter Plot of log_price vs. Accommodates",
       subtitle = "Filtered by Property Types: Apartment, Condominium, House",
       x = "log_price",
       y = "Accommodates") +
  scale_color_manual(values = c("Apartment" = "blue", "Condominium" = "green", "House" = "red")) +
  theme_minimal() +
  theme(legend.title = element_blank(),
        legend.position = "bottom")

##############################

top_zipcodes <- filtered_data %>%
  group_by(zipcode) %>%
  summarise(Count = n()) %>%
  top_n(10, Count) %>%
  pull(zipcode)

filtered_data_top_zipcodes <- filtered_data %>%
  filter(zipcode %in% top_zipcodes)

filtered_data_top_zipcodes$zipcode <- factor(filtered_data_top_zipcodes$zipcode, levels = top_zipcodes)

ggplot(filtered_data_top_zipcodes, aes(x = zipcode, y = combined_score)) +
  geom_boxplot(fill = "skyblue", outlier.color = "red", outlier.shape = 1) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Box Plot of Combined Score by Zipcode",
       x = "Zipcode",
       y = "Combined Score")

###############################

filtered_data <- filtered_data %>% 
  filter(!is.na(bathrooms) & !is.na(combined_score))

ggplot(filtered_data, aes(x = bathrooms, y = combined_score)) +
  geom_point(alpha = 0.6, color = "blue") + 
  theme_minimal() +
  labs(title = "Scatter Plot of Combined Score vs. Bathrooms",
       x = "Number of Bathrooms",
       y = "Combined Score")

#####

filtered_data <- filtered_data %>%
  filter(!is.na(bedrooms) & !is.na(combined_score))

ggplot(filtered_data, aes(x = bedrooms, y = combined_score)) +
  geom_point(alpha = 0.6, color = "green") +
  theme_minimal() +
  labs(title = "Scatter Plot of Combined Score vs. Bedrooms",
       x = "Number of Bedrooms",
       y = "Combined Score")

#############

library(dplyr)
library(ggplot2)

if(all(c("log_price", "property_type", "accommodates", "bathrooms", "bed_type", "bedrooms", "combined_score") %in% names(filtered_data))) {
  
  filtered_data$property_type <- as.factor(filtered_data$property_type)
  filtered_data$bed_type <- as.factor(filtered_data$bed_type)
  
  model <- lm(combined_score ~ log_price + property_type + accommodates + bathrooms + bed_type + bedrooms, 
              data = filtered_data)
  
  summary_model <- summary(model)
  print(summary_model)
} else {
  cat("Error: Not all required variables are present in the dataset.\n")
}
```

## Results

Loft is the best property tyoe, however because its so rare to find I think its better to declare appartments and houses as the best property type.

The loft property type is very unavaliable and not many exist on the airbnb market. However, if you do find one they are heavily rented,
liked, and have a good price.

Private rooms and entire homes are the best room type. Avoid shared rooms as they are not as rented and are very low in price. Private rooms are okay if you have low capital and need an initial investment property. Entire homes are optimal for maximizing combined score and price.

the distribution of the accomodates variable is normalish. Its very hit or miss as the accomodates rise. AirBnB's appear to do better when they are tailored to tourist needs witch is most often couples, singles, or a few friends. More accomodates does not coorelate to better combined score or price. Its better to have 2-3 bathrooms or bedrooms and 8 accomodates for the best results.

Zipcodes show to favor the ones that are in the northern part of chicago, near Old Town. The more central/north located properties do better in renting frequency, price, and reviews.





## Conclusion

In conclusion my analysis shows that in Chicago Homes / appartments rthat are located in the northern part of the city and include the entire home/appartment in the rental and additionally have 8 accomodates, 1 - 2 bathrooms, 1 - 3 bedrooms do better in renting frequency, price, and rent reviews. This is because AirBnB reqrds the 
